---
title: "Usage of PPoS package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{usecase_ppos}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#### Install the package.

```{r setup}
library(ppos)
```

# PPoS calculations 

  - PPoS given historical data on the same endpoint
      - Using MAP (Meta-analystic predictive prior) Approach : **ppos_MAP()**
      
      
  - PPoS given data from a surrogate endpoint **ppos_surrogate2()**

## 1. PPoS calculations: PPoS given historical data on the same endpoint

#### PPoS calculation at the planning stage of phase 3 study

Given data ($Y_{obs}$) from the historical trials of similar drugs, we calculate the PPoS for the future endpoint $Y^*$.
Import data from 27 historical studies and define user inputs such as estimated standard error of the future endpoint $\sigma_*= 0.1$, alpha and hypothesis $H_0: \theta =0$ vs $H_1: \theta < 0$, where $\theta=log(HR)$. 



```{r}
library(readr)



ICBdemo_univariateOS <- read_csv("data/ICBdemo_univariateOS.csv",show_col_types = FALSE)
df <- ICBdemo_univariateOS
Hy.in <- df$logpfshr 
Hy.se.in <- df$sd_lpfsHR1 
input<- NULL

# -------- User Inputs
input$sigma.fs2 <- 0.1  # this is the estimate SE for the future outcome
input$alternative2 <- "less"
input$alpha2 <- 0.05

# --------  # -------- # --------
# -------- fit MAP model
# --------  # -------- # --------

out.jags <- ppos_MAP(Hy=Hy.in,
                     Hy.se=Hy.se.in,
                     tau.dist=NULL,
                     tau.prior=NULL,
                     alpha=as.numeric(input$alpha2),
                     sigma.fs=input$sigma.fs2,
                     alternative= input$alternative2 ,
                     n.burnin=4000, n.iter=10000
)

ppos(out.jags)

```
```{r}
forestplot3(out.jags)
```




## 2. PPoS calculations: PPoS given data from a surrogate endpoint

Given the information on the surrogate endpoint (PFS), in the ongoing study, we calculate the PPoS for the yet to be observed final endpoint OS. Observed treatment effect is $exp(-0.36)=0.7$ i.e., a PFS HR=0.7 was observed, with an estimated standard error of 0.1. Let us consider a standard error of $\sigma_* =0.1$ for the future endpoint. We consider a HalfNormal$(0,0.65)$ prior on the between-study correlation coefficient $\rho$ with a constant within- study correlation $r=0.5$ across all studies. Similarly $r$ can be considered as any other value estimated from the given data. How to estimate $r$ is discussed in our manuscript.

```{r}
ICBdemo_bivariate <- read_csv("data/ICBdemo_bivariate.csv",show_col_types = FALSE)
head(ICBdemo_bivariate)
df <- ICBdemo_bivariate
Hy.in <- cbind(df$logoshr,df$logpfshr)
Hy.se.in <- cbind(df$sd_lOSHR1,df$sd_lpfsHR1)

```

```{r}
out <- ppos_surrogate2(Hy=Hy.in,
                     Hy.se=Hy.se.in,
                     rho.hat=0.5,
                     tau.dist="HalfNormal", # Uniform, InvGamma
                     tau.prior=c(0,0.5),
                     alpha=0.05,
                     X.cs= -0.36, # This can be varied over a sequence
                     sigma.fs=0.1,
                     sigma.cs=0.1,
                     alternative = c("less"),
                     n.burnin=4000, n.iter=10000)
ppos(out)
res <- data.frame(c("rho","Y.star"),round(c(out["rho",1],out["Tf",1]),4))
colnames(res) <- c("","")
res
```
The estimated between-study correlation is 0.89 and the predicted overall survival HR = 0.69.

